---
title: "CFerrari_Assignment10"
author: "Charley Ferrari"
date: "Friday, April 03, 2015"
output: pdf_document
---

Problem Set 1
============

You'll verify for yourself that PageRank works by performing calculations on a small
universe of web pages.Let's use the 6 page universe that we had in the course notes. 
For this directed graph, perform the following calculations in R.


Form the A matrix. Then, introduce decay and form the B matrix as we did in
the course notes.

```{r}
A <- matrix(c(0, 1/2, 1/2, 0, 0,  0,
             0, 0, 1, 0, 0, 0,
             1/4, 1/4, 0, 0, 1/4, 1/4,
             0, 0, 0, 0, 1/2, 1/2,
             0, 0, 0, 1/2, 0, 1/2,
             0, 0, 1/2, 1/2, 0, 0), nrow=6)

# Introduce decay d = 0.85

d <- 0.85

B <- 0.85*A + (0.15/6)


```

Start with a uniform rank vector r and perform power iterations on B till convergence.
That is, compute the solution $r = B^k × r$. Attempt this for a sufficiently
large k so that r actually converges

```{r}

# Technically, this function will multiply A by B 'power' times
# But I'm going to use this to calculate powers.

matrixPower <- function(A, power, B){
  if(power == 1){
    return(A)
  }else{
    return(matrixPower(A %*% B, power - 1, B))
  }
}

r <- matrix(c(1/6,1/6,1/6,1/6,1/6,1/6),nrow=6)

k <- 500

rk <- matrixPower(B,k,B) %*% r

rk




```

Compute the eigen-decomposition of B and verify that you indeed get an eigenvalue
of 1 as the largest eigenvalue and that its corresponding eigenvector is the same
vector that you obtained in the previous power iteration method. Further, this
eigenvector has all positive entries and it sums to 1.

```{r}

eigen(B)$values[1]

# Need to scale the Eigenvectors, I just divided the vectors to get my scale.

(rk / eigen(B)$vectors[,1]) * eigen(B)$vectors[,1]




```

Use the graph package in R and its page.rank method to compute the Page Rank
of the graph as given in A. Note that you don't need to apply decay. The package
starts with a connected graph and applies decay internally. Verify that you do get
the same PageRank vector as the two approaches above

```{r}

library(igraph)


relations <- data.frame(from=c(1,1,2,3,3,3,3,4,4,5,5,6),
                        to = c(2,3,3,1,2,5,6,5,6,4,6,4),
                        weight = c(1,1,1,1,1,1,1,1,1,1,1,1))


g <- graph.data.frame(relations, directed=TRUE)

page.rank(g)

# Once again to scale: 

(rk / page.rank(g)$vector) * page.rank(g)$vector

plot(g)

```


