---
title: "charleyferrari_week5hw"
author: "Charley Ferrari"
date: "October 24, 2015"
output: pdf_document
---

### 5.6: Working Backwards, Part II ###

A 90% confidence interval for a population mean is (65,77). The population 
distribution is approximately normal and the population standard deviation is 
unknown.  This confidence interval is based on a simple random sample of 25 
observations. Calculate the sample mean, the margin of error, and the sample 
standard deviation.

We can solve backwards from knowing what a confidence interval is:

$$ CI = \bar{x} \pm t_{24}^{*} SE $$

$$ 65 = \bar{x} - t_{24}^{*} SE $$

```{r}

xbar <- mean(65,77)

moe <- xbar-65

t <- qt(0.975,24)

SE <- (xbar-65)/t

xbar
moe

```

$$ SE = \sqrt{\frac{\sigma}{n}} $$

```{r}

n <- 25

sd <- n*SE^2

sd

```


### 5.14: SAT Scores ###

SAT scores of students at an Ivy League college are distributed with a standard 
deviation of 250 points. Two statistics students, Raina and Luke, want to 
estimate the average SAT score of students at this college as part of a class 
project. They want their margin of error to be no more than 25 points.

a. Raina wants to use a 90% confidence interval. How large a sample should she 
collect?

$$ MOE = t \times \sqrt{\frac{\sigma}{n}} $$

After trying to calculate this with z, it seems Raina would be able to acheive 
her margin of error using less than 30 observations. Since the t statistic 
changes depending on the number of observations, we can't just solve this 
algebraically. I'll calculate a few versions of the margin of error:

```{r}

n <- seq(2,25,by=1)
t <- qt(0.95,n-1)
sd <- 250
se <- sqrt(sd/n)
moe <- t*se

moe

```


It looks like the margin of error can be acheived with a sample of 4 (although 
it probably wouldn't be safe to use this small a sample size.)

### 5.20, High School and Beyond, Part I ###

The National Center of Education Statistics conducted a survey of high school 
seniors, collecting test data on reading, writing, and several other subjects.
Here we examine a simple random sample of 200 students from this survey. Side-
by-side box plots of reading and writing scores as well as a histogram of the 
differences in scores are shown below.

a. Is there a clear difference in the average reading and writing scores? 

There is a clear difference according to the box plots. The average reading 
score is around 50, while the average writing score is around 55.

b. Are the reading and writing scores of each student independent of each other?
They are not, because this data is paired. Each student has a reading and a 
writing score.

c.  Create hypotheses appropriate for the following research question: Is there 
an evident difference in the average scores of students in the reading and 
writing exam?

$H_0$: $\mu_{reading} = \mu_{writing}$, or $\mu_{reading} - \mu_{writing} = 0$
$H_a$: $\mu_{reading} \neq \mu_{writing}$ or $\mu_{reading} - \mu_{writing} \neq 0$

d. Check the conditions required to complete this test.

Our sample size is large enough at 200, and the differences appears to be 
relatively normally distributed. We should be able to use z-scores and the 
normal distribution to perform our inference, but we can use the t-distribution 
just to be sure.

e. The average observed difference in scores is $\bar{x}_{read-write} = -0.545$, 
and the standard deviation of the differences is 8.887 points. Do these data 
provide convincing evidence of a difference between the average scores on the 
two exams?

We're using the second form of the hypotheses I stated in part c.

$$ T = \frac{\bar{x} - 0}{\sigma / \sqrt{n}} $$

```{r}

xbar <- -0.545
sd <- 8.887
n <- 200
t <- xbar/(sd/sqrt(n))

pt(t,n-1)

```

The p-value isn't below 0.05, so we can't reject the null hypothesis that the 
averages of the reading and writing scores are the same.

f. What type of error might we have made? Explain what the error means in the 
context of the application.

We might have made a type II error in this case, since we have failed to reject 
the null hypothesis. This is a false negative, failing to detect a difference 
that is present.

g. Based on the results of this hypothesis test, would you expect a confidence 
interval for the average difference between the reading and writing scores to 
include 0? Explain your reasoning.

I would expect the confidence interval to include 0 because we failed to reject 
the null hypothesis. Finding a p-value above 0.05 is the same as finding a value 
within the 95% confidence interval.

### Fuel efficiency of manual and automatic cars, Part I ###

Each year the EPA releases fuel economy data on cars manufactured in that year. 
Below are summary statistics on fuuel efficiency (in miles/gallon) from random 
samples of cars with manual and automatic transmissions manufactured in 2012. Do 
these data provide strong evidence of a difference between the average fuel 
efficiency of cars with manual and automatic transmissions in terms of their 
average city mileage? Assume that conditions for inference are satisfied.

$H_0$: $\bar{x}_{auto} - \bar{x}_{manual} = 0$
$H_a$: $\bar{x}_{auto} - \bar{x}_{manual} \neq 0$

```{r}

xbarauto <- 16.12
xbarman <- 19.85
sdauto <- 3.58
sdman <- 4.51
nauto <- 26
nman <- 26
se <- sqrt(sdauto^2/nauto + sdman^2/nman)
tstat <- (xbarauto-xbarman)/se

pt(tstat,nauto-1)*2

```

the p-value of our t-stat (multiplied by 2 since this is a two-tailed test) is 
less than 0.05, so we can reject the null hypothesis. It would appear that the 
average fuel economy of automatic transmissions is significantly less than the 
average fuel economy of manual transmissions.

### 5.48: Work hours and education ###

The General Social Survey collects data on demographics, education, and work, 
among many other characteristics of US residents. Using ANOVA, we can consider 
educational attainment levels for all 1,172 respondents at once. Below are the 
distributions of hours worked by educational attainment and relevant summary 
statistics that will be helpful in carrying out this analysis.

a. Write hypotheses for evalulating whether the average number of hours worked 
varies across the five groups.

$H_0$: $\mu_{<hs} = \mu_{hs} = \mu_{jc} = \mu_{b} = \mu_{g}$
$H_a$: at least one of these means are different.

b. Check conditions and describe any assumptions you must make to proceed with 
the test.


The observations must be independent. This is a representative survey of US 
residents that's below 10% of the population, so as long as this data is truly 
a random sample this condition should hold.

The data must also be approximately normal. The data appears normal enough from 
the boxplots, and for a sample size of around 1000 this should be usable.

The data must also have comparable standard deviations. This condition also 
appears to be satisfied looking at the table, with the potential outlier of 
junior college with a standard deviation of 18.1.

c. Below is part of the output associated with this test. Fill in the empty 
cells.

Lets work our way backwards. First, lets calculate our two degrees of freedom:

```{r}
k <- 5
n <- 1172
df1 <- k-1
df2 <- n-k
df1
df2

```

Now that we know the degrees of freedom, we can use the qf function on the 
Pr(>F) to get the F-statistic:

```{r}

prf <- 0.0682
fstat <- qf(1-prf,df1,df2)
fstat

```

The f-stat is equal to MSG/MSE. We have MST, so we can use that to get MSE:

```{r}

MSG <- 501.54
MSE <- MSG/fstat

```

$MSG = \frac{1}{df_g} SSG$, so we can just multiply MSG by $df_g$ to get SSG:

```{r}

SSG <- df1*MSG
SSE <- df2*MSE
SSG
SSE

```

Just to test it out, I calculated SSE in this way as well. It seems to check out!

SST = SSG + SSE, and $df_{total} = df_g + df_e$

```{r}

SST <- SSG + SSE
dft <- df1 + df2
SST
dft
```

d. What is the conclusion of this test?

the p-value is somewhat greater than 0.05, so we do not reject the null 
hypothesis that the five means are equal to eachother.
