---
title: "Multiple linear regression"
output:
  pdf_document: default
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
---

## Grading the professor

Many college courses conclude by giving students the opportunity to evaluate 
the course and the instructor anonymously. However, the use of these student 
evaluations as an indicator of course quality and teaching effectiveness is 
often criticized because these measures may reflect the influence of 
non-teaching related characteristics, such as the physical appearance of the 
instructor. The article titled, "Beauty in the classroom: instructors' 
pulchritude and putative pedagogical productivity" (Hamermesh and Parker, 2005) 
found that instructors who are viewed to be better looking receive higher 
instructional ratings. (Daniel S. Hamermesh, Amy Parker, Beauty in the 
classroom: instructors pulchritude and  putative pedagogical productivity, 
*Economics of Education Review*, Volume 24, Issue 4, August 2005, Pages 369-376, 
ISSN 0272-7757, 10.1016/j.econedurev.2004.07.013. [http://www.sciencedirect.com/science/article/pii/S0272775704001165](http://www.sciencedirect.com/science/article/pii/S0272775704001165).)

In this lab we will analyze the data from this study in order to learn what goes 
into a positive professor evaluation.

## The data

The data were gathered from end of semester student evaluations for a large 
sample of professors from the University of Texas at Austin. In addition, six 
students rated the professors' physical appearance. (This is aslightly modified 
version of the original data set that was released as part of the replication 
data for *Data Analysis Using Regression and Multilevel/Hierarchical Models* 
(Gelman and Hill, 2007).) The result is a data frame where each row contains a 
different course and columns represent variables about the courses and professors.

```{r load-data}
load("more/evals.RData")
```

variable         | description
---------------- | -----------
`score`          | average professor evaluation score: (1) very unsatisfactory - (5) excellent.
`rank`           | rank of professor: teaching, tenure track, tenured.
`ethnicity`      | ethnicity of professor: not minority, minority.
`gender`         | gender of professor: female, male.
`language`       | language of school where professor received education: english or non-english.
`age`            | age of professor.
`cls_perc_eval`  | percent of students in class who completed evaluation.
`cls_did_eval`   | number of students in class who completed evaluation.
`cls_students`   | total number of students in class.
`cls_level`      | class level: lower, upper.
`cls_profs`      | number of professors teaching sections in course in sample: single, multiple.
`cls_credits`    | number of credits of class: one credit (lab, PE, etc.), multi credit.
`bty_f1lower`    | beauty rating of professor from lower level female: (1) lowest - (10) highest.
`bty_f1upper`    | beauty rating of professor from upper level female: (1) lowest - (10) highest.
`bty_f2upper`    | beauty rating of professor from second upper level female: (1) lowest - (10) highest.
`bty_m1lower`    | beauty rating of professor from lower level male: (1) lowest - (10) highest.
`bty_m1upper`    | beauty rating of professor from upper level male: (1) lowest - (10) highest.
`bty_m2upper`    | beauty rating of professor from second upper level male: (1) lowest - (10) highest.
`bty_avg`        | average beauty rating of professor.
`pic_outfit`     | outfit of professor in picture: not formal, formal.
`pic_color`      | color of professor's picture: color, black & white.

## Exploring the data

1.  Is this an observational study or an experiment? The original research
    question posed in the paper is whether beauty leads directly to the
    differences in course evaluations. Given the study design, is it possible to
    answer this question as it is phrased? If not, rephrase the question.
    
    This is an observational study since we're observing whether or not a 
    professor's looks correlates with ratings.

2.  Describe the distribution of `score`. Is the distribution skewed? What does 
    that tell you about how students rate courses? Is this what you expected to 
    see? Why, or why not?
    
```{r}

library(ggplot2)

ggplot(evals,aes(x=score)) + geom_histogram()


```
    
    It looks like the distribution of scores is negatively skewed, with more 
    students giving higher ratings than lower ones. I actually expected to see 
    a positive skew for scores, with students who were upset with a course more 
    likely to give a (more negative) grade.
    

3.  Excluding `score`, select two other variables and describe their relationship 
    using an appropriate visualization (scatterplot, side-by-side boxplots, or 
    mosaic plot).
    
```{r}

library(vcd)

mosaic(~ ethnicity + gender, data=evals)

```
    
    Using a mosaic plot to compare professors by ethnicity and gender, it seems 
    like it is more likely for minority professors to be women than non-minority 
    professors.

## Simple linear regression

The fundamental phenomenon suggested by the study is that better looking teachers
are evaluated more favorably. Let's create a scatterplot to see if this appears 
to be the case:

```{r scatter-score-bty_avg}
plot(evals$score ~ evals$bty_avg)
```


Before we draw conclusions about the trend, compare the number of observations 
in the data frame with the approximate number of points on the scatterplot. 
Is anything awry?

4.  Replot the scatterplot, but this time use the function `jitter()` on the
    $y$- or the $x$-coordinate. (Use `?jitter` to learn more.) What was
    misleading about the initial scatterplot?
    
```{r}

ggplot(evals,aes(x=bty_avg, y=score)) + geom_point(position = "jitter")

```
    
    It looks like there were points that had the same value. When viewed in 
    a scatter plot without jittering, the points would appear to be exactly 
    the same. By jittering, we can see clusters of points that were otherwise 
    hidden in this way.
    


5.  Let's see if the apparent trend in the plot is something more than
    natural variation. Fit a linear model called `m_bty` to predict average
    professor score by average beauty rating and add the line to your plot
    using `abline(m_bty)`. Write out the equation for the linear model and
    interpret the slope. Is average beauty score a statistically significant
    predictor? Does it appear to be a practically significant predictor?
    
```{r}

m_bty <- lm(score ~ bty_avg, data=evals)

summary(m_bty)

ggplot(evals,aes(x=bty_avg, y=score)) + geom_point(position = "jitter") + 
  stat_smooth(method="lm")

```
    
    The equation for this line is: score = 3.88034 + 0.06664 $\times$ bty_avg
    
    It appears to be somewhat predictive, but there also appear to be lots of 
    residuals. Our R-squared tells us the same thing: at only 0.03502, this 
    model isn't very strong.
    


6.  Use residual plots to evaluate whether the conditions of least squares
    regression are reasonable. Provide plots and comments for each one (see
    the Simple Regression Lab for a reminder of how to make these).
    
```{r}

hist(m_bty$residuals)

```
    
    Linearity: This condition appears to be partially met, there is a wide 
    variance, but there is a certain degree of linearity.
    
    Nearly normal residuals: This condition seems to be somewhat met... The 
    data iss a bit negatively skewed however.
    
    Constant variability: This condition, according to the scatter plot, might 
    not be met. First of all, there is a chunk of scores between 4 and 5 that 
    might be skewing this condition. Higher beauty scores are more sparse, and 
    the score cap at 5 seems to be affecting the variability as the beauty score 
    gets higher.

## Multiple linear regression

The data set contains several variables on the beauty score of the professor: 
individual ratings from each of the six students who were asked to score the 
physical appearance of the professors and the average of these six scores. Let's 
take a look at the relationship between one of these scores and the average 
beauty score.

```{r bty-rel}
plot(evals$bty_avg ~ evals$bty_f1lower)
cor(evals$bty_avg, evals$bty_f1lower)
```

As expected the relationship is quite strong - after all, the average score is 
calculated using the individual scores. We can actually take a look at the 
relationships between all beauty variables (columns 13 through 19) using the 
following command:

```{r bty-rels, eval = FALSE}
plot(evals[,13:19])
```

These variables are collinear (correlated), and adding more than one of these 
variables to the model would not add much value to the model. In this 
application and with these highly-correlated predictors, it is reasonable to use
the average beauty score as the single representative of these variables.

In order to see if beauty is still a significant predictor of professor score 
after we've accounted for the gender of the professor, we can add the gender 
term into the model.

```{r scatter-score-bty_avg_gender}
m_bty_gen <- lm(score ~ bty_avg + gender, data = evals)
summary(m_bty_gen)
```

7.  P-values and parameter estimates should only be trusted if the
    conditions for the regression are reasonable. Verify that the conditions
    for this model are reasonable using diagnostic plots.
    
```{r}

ggplot(evals,aes(x=gender,y=score)) + geom_boxplot()

```
    
    Here we're dealing with a categorical variable, so it made more sense to 
    view these as boxplots. It does seem like there is a difference between 
    male and female scores.
    
```{r}

hist(m_bty_gen$residuals)

```
    
    The residuals seem to be fairly normal as well, if a bit skewed once again.
    

8.  Is `bty_avg` still a significant predictor of `score`? Has the addition
    of `gender` to the model changed the parameter estimate for `bty_avg`?
    
```{r}
summary(m_bty_gen)

```
    
    It does still seem like average beauty is a significant predictor of 
    score, and the gender of the student has only improved our $R^2$ a bit 
    from 0.03 to 0.05 (still not a great $R^2$)
    
    

Note that the estimate for `gender` is now called `gendermale`. You'll see this 
name change whenever you introduce a categorical variable. The reason is that R 
recodes `gender` from having the values of `female` and `male` to being an 
indicator variable called `gendermale` that takes a value of $0$ for females and
a value of $1$ for males. (Such variables are often referred to as "dummy" 
variables.)

As a result, for females, the parameter estimate is multiplied by zero, leaving 
the intercept and slope form familiar from simple regression.

\[
  \begin{aligned}
\widehat{score} &= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg + \hat{\beta}_2 \times (0) \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg\end{aligned}
\]

We can plot this line and the line corresponding to males with the following 
custom function.

```{r twoLines, eval = FALSE}
multiLines(m_bty_gen)
```

9.  What is the equation of the line corresponding to males? (*Hint:* For
    males, the parameter estimate is multiplied by 1.) For two professors
    who received the same beauty rating, which gender tends to have the
    higher course evaluation score?
    
    For males, the equation of the line is: 
    score = 3.91973 + 0.07416 $\times$ bty_avg
    
    This is just the genderemale parameter plus the intercept, since if we're 
    only looking at males this will always be 1. For two professors who 
    received the same beauty rating, males would be predicted to have higher 
    scores.

The decision to call the indicator variable `gendermale` instead of`genderfemale`
has no deeper meaning. R simply codes the category that comes first 
alphabetically as a $0$. (You can change the reference level of a categorical 
variable, which is the level that is coded as a 0, using the`relevel` function. 
Use `?relevel` to learn more.)

10. Create a new model called `m_bty_rank` with `gender` removed and `rank` 
    added in. How does R appear to handle categorical variables that have more 
    than two levels? Note that the rank variable has three levels: `teaching`, 
    `tenure track`, `tenured`.
    
```{r}
m_bty_rank <- lm(score ~ bty_avg + rank, data=evals)
summary(m_bty_rank)
levels(evals$rank)

```
    
    If there are n levels, R adds n-1 variables. These represent n-1 categories, 
    with the nth category represented by all the variables being 0.

The interpretation of the coefficients in multiple regression is slightly 
different from that of simple regression. The estimate for `bty_avg` reflects
how much higher a group of professors is expected to score if they have a beauty
rating that is one point higher *while holding all other variables constant*. In
this case, that translates into considering only professors of the same rank 
with `bty_avg` scores that are one point apart.

## The search for the best model

We will start with a full model that predicts professor score based on rank, 
ethnicity, gender, language of the university where they got their degree, age, 
proportion of students that filled out evaluations, class size, course level, 
number of professors, number of credits, average beauty rating, outfit, and 
picture color.

11. Which variable would you expect to have the highest p-value in this model? 
    Why? *Hint:* Think about which variable would you expect to not have any 
    association with the professor score.
    
    This is actually a toss up for me... All of these variables could 
    conceptually affect student rankings. If I had to guess, I would think 
    that ethnicity might play the lowest role.

Let's run the model...

```{r m_full, eval = FALSE, tidy = FALSE}
m_full <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full)
```

12. Check your suspicions from the previous exercise. Include the model output
    in your response.
    
    Ethnicty played a low role in accordance to what I guessed. According to 
    the p-values, other variables that didn't predict too well were stats about 
    the class, and tenure (which was actually my second guess.)

13. Interpret the coefficient associated with the ethnicity variable.

As mentioned above, ethnicity does not appear to be significant in predicting 
the scores of professors. The p-value here is 0.12, which is above our 5% 
threshold.

14. Drop the variable with the highest p-value and re-fit the model. Did the
    coefficients and significance of the other explanatory variables change?
    (One of the things that makes multiple regression interesting is that
    coefficient estimates depend on the other variables that are included in
    the model.) If not, what does this say about whether or not the dropped
    variable was collinear with the other explanatory variables?
    
```{r}
m_full_nosingle <- lm(score ~ rank + ethnicity + gender + language + age +
                        cls_perc_eval + cls_students + cls_level +
                        cls_credits + bty_avg + pic_outfit + pic_color, 
                      data = evals)
summary(m_full_nosingle)

```
    
    Our R squared adjusted changed slightly, but more importantly, our 
    p-values seemed to have changed. It seems like now we have no p-value 
    above 0.3, while before we had some up to 0.7.

15. Using backward-selection and p-value as the selection criterion,
    determine the best model. You do not need to show all steps in your
    answer, just the output for the final model. Also, write out the linear
    model for predicting score based on the final model you settle on.
    
    The model created above actually seems to be best. I tried to remove the 
    next highest p-value, cls_level, but the R-squared fell slightly, and a 
    few of the p-values were higher than in the previous model.

16. Verify that the conditions for this model are reasonable using diagnostic 
    plots.
    
```{r}

hist(m_full_nosingle$residuals)

```

It seems like the residuals are as normal as in the models we were studying 
before. If we wanted to truly check all diagnostic plots, we'd want to check the 
scatter plots to make sure our data looks fairly linear

17. The original paper describes how these data were gathered by taking a
    sample of professors from the University of Texas at Austin and including 
    all courses that they have taught. Considering that each row represents a 
    course, could this new information have an impact on any of the conditions 
    of linear regression?
    
    I'd be interested to see if they double count professors in classes. In 
    order to properly model this, they should link the scores given to 
    professors in different classes, and average them (making sure to account
    for the increased sample)

18. Based on your final model, describe the characteristics of a professor and 
    course at University of Texas at Austin that would be associated with a high
    evaluation score.
    
    Based on the final model, the qualities being associated with a high 
    evaluation score are being of teaching rank, non-minority, male, speaking 
    english, younger, and teaching one credit classes.

19. Would you be comfortable generalizing your conclusions to apply to professors
    generally (at any university)? Why or why not?
    
    It seems like this might be fairly generalizeable. Personal qualities of the 
    professor, such as age, language, gender, and ethnicity might be 

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). This lab was written by 
Mine &Ccedil;etinkaya-Rundel and Andrew Bray.
</div>
