---
title: "Panel Data World Bank"
author: "Charley Ferrari"
date: "May 6, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Exploratory analysis of Tuberculosis Incidence

- Facet plot of Rates for different countries over time:

```{r facet, warning=FALSE, message=FALSE}

library(WDI)
library(reshape2)
library(ggplot2)
library(ggthemes)
library(psychometric)
library(plm)
library(dplyr)

wb <- WDI(country="all", indicator=c("DT.ODA.ODAT.PC.ZS","SH.TBS.INCD"),
          start=2005, end=2015, extra=TRUE)
wb <- filter(wb, income == 'Low income')

wbcastODA <- dcast(country ~ year, data=wb, value.var = 'DT.ODA.ODAT.PC.ZS')

wbcastODA <- wbcastODA[!is.na(wbcastODA[,'2005']),]

wbcastODA$country <- factor(wbcastODA$country)

countrylist <- levels(wbcastODA$country)

wbcastTUB <- dcast(country ~ year, data=wb, value.var='SH.TBS.INCD')

wb <- wb %>%
  filter(country %in% countrylist)

wbmodel <- wb %>%
  filter(year %in% 2005:2014) %>%
  dplyr::select(country, year, DT.ODA.ODAT.PC.ZS, SH.TBS.INCD)

ggplot(wbmodel, aes(x=year, y=SH.TBS.INCD)) + geom_line() + 
  facet_wrap( ~ country) + theme_tufte()

```

## Exploratory analysis of Tuberculosis Incidence

- There appears to be low variation within most groups, but most 

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

