---
title: "is604hw4test"
author: "Charley Ferrari"
date: "October 17, 2015"
output: pdf_document
---
#### e. Importance Sampling ###

To pick a function here, I wanted to use an upside-down parabola that crossed 
the x-axis at -5 and 5, which would be $\frac{-3(x-5)(x+5)}{500}$

This would be difficult to solve for an inverse transform, so lets sample from 
it using acceptance/rejection.

The max of my parabola is at 0.15, and cofx(0,1) (the normal distribution's 
maximum) is 0.3989. So, I can safely say:

$$ \frac{f(x)}{g(x)} \leq \frac{0.15}{0.1} \leq 1.5 $$


So, to sample from f(x), I'll generate y from g(x), generate a random u from 
Uniform(0,1), and accept y if $U < \frac{f(y)}{cg(y)}$

The sampling would work like this:

$$ \frac{f(x)}{cg(x)} = \frac{-3(x-5)(x+5)}{500} \times \frac{10}{1.5} = \frac{(x-5)(x+5)}{25} $$

```{r}

rcharleyparabola <- function(n){
  n <- 1000
  k <- 0
  j <- 0
  y <- numeric(n)
  
  while(k < n){
    u <- runif(1)
    j <- j+1
    x <- runif(1,min=-5,max=5)
    if(-(x-5)*(x+5)/25 > u){
      k <- k+1
      y[k] <- x
    }
  }
  
  return(y)
  
}
```

So lets perform our importance sampling:

```{r}

isfunction <- function(n,D){
  m <- n*D
  x <- matrix(rcharleyparabola(m),nrow=n)
  fg <- apply(x,1,cofx,D=D) / (-3*(x-5)*(x+5)/500)
  
  return(fg)
}

iscreate <- function(D){
  
  n <- seq(1000,10000,by=1000)
  means <- c()
  sds <- c()
  coefvars <- c()
  
  for(i in n){
    thetasample <- replicate(100,atfunction(i,D))
    means <- c(means, mean(thetasample))
    sds <- c(sds, sd(thetasample))
    coefvars <- c(coefvars, mean(thetasample)/sd(thetasample))
  }
  
  table <- data.frame(n=n, means=means, sds=sds, coefvars=coefvars, D=D,
                      method="importancesampling")
  
  return(table)
  
}

```

