---
title: "IS604_charleyferrari_hw4"
author: "Charley Ferrari"
date: "October 12, 2015"
output: pdf_document
---

Monte Carlo test:

```{r}

m <- 10000
x <- runif(m, min=2, max=4)
theta.hat <- mean(exp(-x))*2
theta.hat
exp(-2) - exp(-4)

```

a. Crude Monte Carlo

For sample sizes 1000 to 10000 (in increments of 1000), obtain 100 estimates 
for E[c(x)] when D = 1, using crude Monte Carlo sampling.

First, lets create a function cofx, that will perform our function. It's meant 
to take in a D-dimensional vector, and return:

$$ c(x) = \frac{1}{(2 \pi)^{\frac{D}{2}}} $$

```{r}

D <- 1

cofx <- function(x,D){
  theta.hat <- (1/((2*pi)^(D/2)))*exp((-1/2)*(t(x) %*% x))
  return(theta.hat)
}

repeatfunction <- function(n,D){
  
  m <- D*n
  x <- matrix(runif(m, min=-5, max=5),nrow=D)
  
  return(mean(apply(x,2,cofx,D=D)))
  
}

n <- seq(1000,10000,by=1000)
means <- c()
sds <- c()
coefvars <- c()

for(i in n){
  thetasample <- replicate(100,repeatfunction(i,D))
  means <- c(means, mean(thetasample))
  sds <- c(sds, sd(thetasample))
  coefvars <- c(coefvars, mean(thetasample)/sd(thetasample))
}

dataunifD1 <- data.frame(n=n, means = means, sds = sds, coefvars = coefvars)

ggplot(dataunifD1, aes(x=n, y=means)) + geom_line() + 
  geom_hline(yintercept=(1/10)^D)
ggplot(dataunifD1, aes(x=n, y=sds)) + geom_line()

```

Lets try this for D = 2

```{r}

D <- 2

means <- c()
sds <- c()
coefvars <- c()

for(i in n){
  thetasample <- replicate(100,repeatfunction(i,D))
  means <- c(means, mean(thetasample))
  sds <- c(sds, sd(thetasample))
  coefvars <- c(coefvars, mean(thetasample)/sd(thetasample))
}

dataunifD2 <- data.frame(n=n, means = means, sds = sds, coefvars = coefvars)

library(ggplot2)

ggplot(dataunifD2, aes(x=n, y=means)) + geom_line() + 
  geom_hline(yintercept=(1/10)^D)
ggplot(dataunifD2, aes(x=n, y=sds)) + geom_line()

```

b. Quasi Random Numbers

```{r}

library(abind)

d <- 1
n <- 100

binvec <- function(n){
  return(rev(as.numeric(intToBits(n))[1:(floor(log(n)/log(2))+1)]))
}

vdc <- function(d,n){
  
  b <- 2
  r <- floor(log(n)/log(b))+1
  
  bitcreate <- function(N,r){
    return(as.numeric(intToBits(N))[1:r])
  }
  
  a <- apply(as.matrix(N,ncol=1),1,bitcreate, r=r)
  bb <- 1/b^(1:r)
  
  return(t(bb) %*% a)
  
}

mcalc <- function(k,r){
  
  polys <- c(1,3,7,11,13,19,25,37,59,47,61,55,41,67,97,91,109,103,
  115,131,193,137,145,143,241,157,185,167,229,171,213,191,
  253,203,211,239,247,285,369,299,425,301,361,333,357,351,
  501,355,397,391,451,463,487)
  
  ivals <- t(matrix(c(1,1,1,1,1,1,1,1,
                    1,3,5,15,17,51,85,255,
                    1,1,7,11,13,61,67,79,
                    1,3,7,5,7,43,49,147,
                    1,1,5,3,15,51,125,141,
                    1,3,1,1,9,59,25,89,
                    1,1,3,7,31,47,109,173,
                    1,3,3,9,9,57,43,43,
                    1,3,7,13,3,35,89,9,
                    1,1,5,11,27,53,69,25,
                    1,3,5,1,15,19,113,115,
                    1,1,7,3,29,51,47,97,
                    1,3,7,7,21,61,55,19,
                    1,1,1,9,23,39,97,97,
                    1,3,3,5,19,33,3,197,
                    1,1,3,13,11,7,37,101,
                    1,1,7,13,25,5,83,255,
                    1,3,5,11,7,11,103,29,
                    1,1,1,3,13,39,27,203,
                    1,3,1,15,17,63,13,65),nrow=8))
  
  m <- ivals[k,]
  ppn <- polys[k]
  c <- tail(binvec(ppn),-1)
  deg <- length(c)
  for(i in 8:r){
    s <- 0
    for(j in 1:deg){
      s <- bitwXor(s,(2^j)*c[j]*m[i-j])
    }
    m <- c(m,bitwXor(s,m[i-deg]))
  }
  
  return(m[1:r])
  
}

vcalc <- function(k,r){
  V <- diag(r)
  if(k>=2){
    for(i in 2:k){
      m <- mcalc(i,r)
      vk <- c()
      for(j in 1:r){
        h <- binvec(m[j])
        h <- c(rep(0,j-length(h)), h, rep(0,r-j))
        vk <- cbind(vk,as.matrix(h,nrow=r))
      }
      V <- abind(V,vk,along=3)
    }
  }
  return(V)
}

sobol <- function(d,n){

  b <- 2
  N <- 0:n
  r <- floor(log(n)/log(b))+1
  bb <- 1/b^(1:r)
  bbb <- rep(bb,n,)
  
  bbb <- bb
  for(i in 1:n){
    bbb <- rbind(bbb, bb)
  }
  rownames(bbb) <- NULL
  
  bitcreate <- function(N,r){
    return(as.numeric(intToBits(N))[1:r])
  }
  
  a <- t(apply(as.matrix(N,ncol=1),1,bitcreate, r=r))
  
  p <- matrix(nrow=n+1,ncol=d)
  
  V <- vcalc(d, r)
  
  for(i in 1:d){
    Vtemp <- V[,,i]
    p[,i] <- bb %*% t(((a%*%t(Vtemp))%%b))
  }
  
  return(p)
  
}

xunif <- runif(100)
yunif <- runif(100)
sobolvar <- sobol(2,100)
xsobol <- sobolvar[,1]
ysobol <- sobolvar[,2]

ggplot(data.frame(x=xunif,y=yunif),aes(x=x,y=y)) + geom_point()
ggplot(data.frame(x=xsobol,y=ysobol),aes(x=x,y=y)) + geom_point()

## Scramble

xsobol <- sobol(1,999)
xunif <- runif(1000)

x <- (xsobol+xunif) %% 1

ysobol <- sobol(1,999)
yunif <- runif(1000)

y <- (ysobol+yunif) %% 1

ggplot(data.frame(x=x,y=y),aes(x=x,y=y))+geom_point()

ggplot(data.frame(x=runif(1000),y=runif(1000)),aes(x=x,y=y))+geom_point()

```

The sobol numbers look much more evenly dispersed, while the uniform random 
numbers are more random. This is even more pronounced when you generate 1000 
samples:

```{r}
xunif <- runif(1000)
yunif <- runif(1000)
sobolvar <- sobol(2,1000)
xsobol <- sobolvar[,1]
ysobol <- sobolvar[,2]

ggplot(data.frame(x=xunif,y=yunif),aes(x=x,y=y)) + geom_point()
ggplot(data.frame(x=xsobol,y=ysobol),aes(x=x,y=y)) + geom_point()

```

Lets run our monte carlo simulation using the sobol numbers:

```{r}
D <- 1

repeatfunction <- function(n,D){
  
  x <- t((sobol(D,n)-0.5)*5)
  
  return(mean(apply(x,2,cofx,D=D)))
  
}

n <- seq(1000,10000,by=1000)
means <- c()
sds <- c()
coefvars <- c()

for(i in n){
  thetasample <- replicate(100,repeatfunction(i,D))
  means <- c(means, mean(thetasample))
  sds <- c(sds, sd(thetasample))
  coefvars <- c(coefvars, mean(thetasample)/sd(thetasample))
}

data1 <- data.frame(n=n, means = means, sds = sds, coefvars = coefvars)

ggplot(data1, aes(x=n, y=means)) + geom_line() + geom_hline(yintercept=(1/10)^D)
ggplot(data1, aes(x=n, y=sds)) + geom_line()


```

c. Antithetic Variates

In my repeatfunction, I set the min and max within the runif function. I could 
have just as easily pulled from a uniform distribution from 0-1, subtracted by
0.5 to center the sample, and then multiply by 10 to get a sample from -5 to 5.

Because my sample is already centered around -5 to 5 however, I wouldn't create 
an antithetic variate by pulling x ~ uniform(0,1) and then finding 1-x. Since 
I'm centered around -, I'll just create my variate from -5 to 5, and then 
create my antithetic variate as -x.

```{r}

D <- 1

repeatfunction <- function(n,D){
  
  m <- D*n
  xhalf1 <- matrix(runif(m/2),nrow=D)
  xhalf2 <- 1-xhalf1
  
  xhalf1 <- (xhalf1*10)-5
  xhalf2 <- (xhalf2*10)-5
  
  fx1 <- apply(xhalf1,2,cofx,D=D)
  fx2 <- apply(xhalf2,2,cofx,D=D)
  fx <- (fx1+fx2)/2
  return(mean(fx))
}

sample <- runif(100/2)
xhalf1 <- matrix(sample,nrow=D)

#repeatfunction <- funcction(n,D){
#  
#  m <- D*n
#  xhalf <- matrix(runif(m/2), nrow=D)
##  x <- cbind(xhalf, 1-xhalf)
#  
#  
#}

n <- seq(1000,10000,by=1000)
means <- c()
sds <- c()
coefvars <- c()

for(i in n){
  thetasample <- replicate(100,repeatfunction(i,D))
  means <- c(means, mean(thetasample))
  sds <- c(sds, sd(thetasample))
  coefvars <- c(coefvars, mean(thetasample)/sd(thetasample))
}

dataantithD1 <- data.frame(n=n, means = means, sds = sds, coefvars = coefvars)

ggplot(dataunifD1, aes(x=n, y=means)) + geom_line() + 
  geom_hline(yintercept=(1/10)^D)
ggplot(dataunifD1, aes(x=n, y=sds)) + geom_line()

dataunifD1$method <- "Crude"
dataantithD1$method <- "Antithetic"

data <- rbind(dataunifD1, dataantithD1)

ggplot(data, aes(x=n, y=means, color=method)) + geom_line() + 
  geom_hline(yintercept=(1/10)^D)
ggplot(data, aes(x=n, y=sds, color=method)) + geom_line()


```



### SCR Problem 6.3 ###

Plot the power curves for the t-test in Example 6.9 for sample sizes 10, 20, 30, 
40, and 50, but omit the standard error bars. Plot the curves on the same graph, 
each in a different color or different
