---
title: "CFerrari_Assignment9"
author: "Charley Ferrari"
date: "Friday, March 27, 2015"
output: pdf_document
---

Problem Set 1
============

This week, we'll empirically verify Central Limit Theorem. We'll write code to run a
small simulation on some distributions and verify that the results match what we expect
from Central Limit Theorem. Please use R markdown to capture all your experiments and
code. Please submit your Rmd file with your name as the filename.

(1) First write a function that will produce a sample of random variable that is distributed
as follows:

$$f(x) = x, 0 \leq x \leq 1$$
$$f(x) = 2, - x 1 < x \leq 2$$

That is, when your function is called, it will return a random variable between 0
and 2 that is distributed according to the above PDF. Please note that this is
not the same as writing a function and sampling uniformly from it. In
the online session this week, I'll cover Sampling techniques. You will
find it useful when you do the assignment for this week. In addition,
as usual, there are one-liners in R that will give you samples from a
function. We'll cover both of these approaches in the online session.

```{r}

library(ggplot2)

invcdf1 <- function(y){
  if (y >= 0 && y <= 2){
    ret <- ifelse(y < 0.5, sqrt(2*y), 2-sqrt(2*(1-y)))
  }
}

pdfcreator <- function(invcdf1){  
  return(sapply(runif(1000),invcdf1))
}

sampleeq1 <- pdfcreator(invcdf1)


```


(2) Now, write a function that will produce a sample of random variable that is distributed
as follows:

$$f(x) = 1 - x, 0 \leq x \leq 1$$
$$f(x) = x - 1, 1 < x \leq 2$$

```{r}

#sqrt(2*x-4)+2

#2 - sqrt(4-2*y)

invcdf2 <- function(y){
  if(y >= 0 && y <= 2){
    ret <- ifelse(y < 0.5, 1-sqrt(1-2*y), 1+sqrt(2*y-1))
  }
}



sampleeq2 <- pdfcreator(invcdf2)




```

(3) Draw 1000 samples (call your function 1000 times each) from each of the above two
distributions and plot the resulting histograms. You should have one histogram for
each PDF. See that it matches your understanding of these PDFs.

The function call is done above, below are the qplots

```{r}

qplot(sampleeq1)

qplot(sampleeq2)

```

(4) Now, write a program that will take a sample set size n as a parameter and the
PDF as the second parameter, and perform 1000 iterations where it samples from
the PDF, each time taking n samples and computes the mean of these n samples.
It then plots a histogram of these 1000 means that it computes

```{r}

qplot(replicate(1000,mean(pdfcreator(invcdf1))))

qplot(replicate(1000,mean(pdfcreator(invcdf2))))

```

(5) Verify that as you set n to something like 10 or 20, each of the two PDFs produce
normally distributed mean of samples, empirically verifying the Central Limit
Theorem. Please play around with various values of n and you'll see that even for
reasonably small sample sizes such as 10, Central Limit Theorem holds.
