---
title: "week4hw609"
author: "Charley Ferrari"
date: "Wednesday, September 16, 2015"
output: pdf_document
---

### Page 191, Question 3 ###

Using Monte Carlo simulation, write an algorthm to calculate an approximation to $\pi$ by 
considering the number of random points selected inside the quarter circle:

$$ Q: x^2 + y^2 = 1,  x \geq 0,  y \geq 0 $$

This is a unit circle, so I'm going to set my M box to be x=1 and y=1. I'll draw random numbers 
from $0 \leq x \leq 1$ and $0 \leq y \leq 1$.

```{r}

montecarlox <- runif(100000, 0, 1)
montecarloy <- runif(100000, 0, 1)
montecarlotest <- montecarlox^2 + montecarloy^2 <= 1

proportion <- length(montecarlotest[montecarlotest])/length(montecarlotest)


```

The area of the bounds set by M is 1, and the proportion of the "hits" that resulted from the Monte 
Carlo test is stated above, along with the area of the quarter circle.

The volume of the whole circle = $\pi r^2$, and from our result above, equals 4 times the area 
we just calculated:

```{r}

calculatedarea <- 4*proportion

```

r=1 in our case here, so what we're left with is an approximation for $\pi$

```{r}

calculatedarea

```

### Page 194, Question 1 ###

Use the middle-square method to generate

a. 10 random numbers using $x_0 = 1009$

```{r}

a <- 1009^2

```

01018081

```{r}

a <- c(a, 180^2)
r <- 180

```

00032400

```{r}

a <- c(a, 324^2)
r <- c(r, 324)

```

00104976

```{r}

a <- c(a, 1049^2)
r <- c(r, 1049)

```

01100401

```{r}

a <- c(a, 1004^2)
r <- c(r, 1004)

```

01008016

```{r}

a <- c(a, 80^2)
r <- c(r, 80)

```

00006400

```{r}

a <- c(a, 64^2)
r <- c(r, 64)

```

00004096

```{r}

a <- c(a, 40^2)
r <- c(r, 40)

```

00001600

```{r}

a <- c(a, 16^2)
r <- c(r, 16)

```

00000256

```{r}

a <- c(a, 2^2)
r <- c(r, 2)

```

00000004

It appears we've ended up at 0.

Going through this process with this seed, and 8 digits seemed to have highlighted a problem. I 
assumed that once smaller numbers began appearing, by random chance I'd be back at higher numbers if 
I continued this algorithm. It would appear that instead there is occassionally a tendency to end up 
at 0. Here's our full list of random numbers with this seed:

```{r}

r

```

### Page 201, Question 4 ###

Construct and perform a Monte Carlo simulation of a horse race.

It appears that the odds add up to more than 1. A quick google search revealed this is simply due to 
the bookie taking a cut:

```{r}
(1/7)+(1/5)+(1/9)+(1/12)+(1/4)+(1/35)+(1/15)+(1/4)

```

So, I'll mutiply each of these percentages by 1/1.13254 to get the underlying probabilities of 
each horse winning:

```{r}

m <- 1/((1/7)+(1/5)+(1/9)+(1/12)+(1/4)+(1/35)+(1/15)+(1/4))

ef <- m/7
ll <- m/5
nl <- m/9
cc <- m/12
pp <- m/4
lh <- m/35
ss <- m/15
dd <- m/4

```

Now I'll compute the cumulative probabilities to make my Monte Carlo simulation easier:

```{r}

ef <- m/7
ll <- m/5 + ef
nl <- m/9 + ll
cc <- m/12 + nl
pp <- m/4 + cc
lh <- m/35 + pp
ss <- m/15 + lh
dd <- m/4 + ss

cumprob <- c(ef,ll,nl,cc,pp,lh,ss,dd)

```

I'll run an iterated simulation, generating random numbers within a for loop. For each random 
number, I'll implement a vectorized solution to find out where along the Cumulative probability 
distribution the number belongs, and hence which horse won.

```{r}

rnlist <- runif(10000,0,1)

horselist <- c()

for(i in rnlist){
  horselist <- c(horselist,sum(i < cumprob))
}

```

In this simulation, the numbers correspond to the horse names in reverse order (8=ef, 7=ll, etc). We
could plot a histogram of the horselist if we wanted to visualize the results:

```{r}

hist(horselist, breaks=8)

```
