---
title: "IS 609 Week 2 hw"
author: "Charley Ferrari"
date: "Friday, September 04, 2015"
output: pdf_document
---

### Page 69, Problem 12 ###

For the scenarios presented in problem 12, identify a problem worth studying and list the variables 
that affect the behavior you have identified. Which variables would be neglected completely? Which 
might be considered as constants initially? Can you identify any submodels you would want to study in 
detail? Identify any data you would want collected.

A company with a fleet of trucks faces increasing maintenance costs as the age and mileage of the 
trucks increase.

Problem worth studying: How can we minimize the total costs of our fleet. Total costs include both 
maintenance costs, and the costs of replacing trucks.

Variables: Number of trucks in the fleet, Replacement cost of a truck, total amount of work required 
of the fleet.
variables for each truck: age, number of hours of planned use, mileage
variables to be neglected: Perhaps we could neglect the type of work. As long as there is a full load,
what matters is how long the truck is on the road, and not where it is going.

Which might be considered as constants initially? Can you identify any submodels you would want to
study in detail?

Initially, we can start with a submodel assuming that the fleet of trucks remain constant. Our model 
at this point is simply modeling the increasing costs of maintaining an aging fleet of vehicles, and 
trying to minimize that cost. The dependent variables would be the number of planned work hours for 
each truck. The total amount of work required of the fleet would be a constraint, and work would be 
allocated among trucks of varying age and mileage.

We can also introduce a longer term model allowing for the replacement of trucks. At some point, 
if maintenance costs are growing without bound, it will make more sense to buy a new truck with lower 
maintenance costs than to continue spending money on an old truck. 

Data I would want collected: I would like to collect data on the maintenance spending for each truck. 
Using this information, I could constantly update my model, and tweak my relationship between age, 
mileage and expected maintenance costs.

### Page 79, Problem 11 ###

In problems 7-12, determine whether the data set supports the stated proportionality model.

$$ y \propto x^3 $$

```{r}

q11data <- data.frame(y=c(0,1,2,6,14,24,37,58,82,114), x=c(1,2,3,4,5,6,7,8,9,10))

q11data$ypred <- q11data$x^3

q11data$k <- q11data$y/q11data$ypred

k <- mean(q11data$k)

kadj <- mean(tail(q11data$k, -1))

q11data$ypredk <- k*(q11data$x^3)

q11data$ypredkadj <- kadj*(q11data$x^3)

q11data

library(reshape2)

library(ggplot2)

ggplot(melt(q11data, id=c("x", "k", "ypred", "ypredk")), aes(x=x, y=value, color=variable)) + 
  geom_point()



```

This data set mostly supports the stated proportionality model. I'd be a bit concerned about the 0 
value for x=1, which throws off the predicted constant k. This might be a situation that is not 
exactly passing through the origin, but can be reasonably approximated by a function that does.

I made my graph based on an adjusted k, where I took the mean of the calculated constant disregarding 
the first 0 value.

### Page 94, Problem 4 ###

Lumber Cutters - Lumber cutters wish to use readily available measurements to estimate the number of 
board feet of lumber in a tree. Assume they measure the diameter of the tree in inches at waist 
height. Develop a model that predicts board feet as a function of diameter in inches.

Use the following data for your test: 

```{r}

q4data <- data.frame(x=c(17, 19, 20, 23, 25, 28, 32, 38, 39, 41), 
                     y=c(19, 25, 32, 57, 71, 113, 123, 252, 259, 294))

```

The variable x is the diameter of a ponderosa poine in inches, and y is the number of board feet 
divided by 10.

a. Consider two separate assumptions, allowing each to lead to a model. Completely analyze each 
model.

i. Assume that all trees are right-circular cylinders, and are approximately the same height.

For this model, I'm going to assume that board feet is proportional to volume. If trees are assumed 
to be approximately the same height, then diameter is going to be the only variable. In this case, 
the area of the circle described by the diameter is what's going to control the board feet. 

The volume of a right cylinder is $\pi r^2 h$, but h is constant. So:

$$ y \propto x^2 $$

Lets test this out:

```{r}

q4data$ypredi <- q4data$x^2

q4data$ki <- q4data$y/q4data$ypredi

ki <- mean(q4data$ki)

q4data$ypredik <- ki*(q4data$x^2)

q4data

ggplot(melt(q4data, id=c("x", "ki", "ypredi")), aes(x=x, y=value, color=variable)) + geom_point()

```

ii. Assume that all trees are right-circular cylinders and that the height of the tree is 
proportional to the diameter.

For this model, a new dimension is added. the volume is still $\pi r^2 h$, but h is proportional to 
the diameter. This new model is now:

$$ y \propto x^3 $$

Lets test this one out:

```{r}

q4data$ypredii <- q4data$x^3

q4data$kii <- q4data$y/q4data$ypredii

kii <- mean(q4data$kii)

q4data$yprediik <- kii*(q4data$x^3)

q4data

ggplot(melt(q4data, id=c("x", "ypredi", "ki", "ypredik", "ypredii", "kii")),
       aes(x=x, y=value, color=variable)) + geom_point()


```

b. Which model appears to be better? Why? Justify your conclusions.

Visually, the plot of the second model seems to match more closely than the plot of the first model. 
It would appear that it is more correct to say height is proportional to diameter, and that 
$y \propto x^3$ rather than $y \propto x^2$.

Intuitively, this makes sense, since taller trees would have fatter trunks.