---
title: "IS621_hw1"
author: "Charley Ferrari"
date: "February 3, 2016"
output: pdf_document
---

```{r}

library(ggplot2)

library(dplyr)

setwd("/Users/Charley/Downloads/cuny/IS 621 Business Analytics and Data Mining/Homework 1")

cigarettes <- read.csv("cigarette-training-data.csv")

pairs(select(cigarettes, -c(State, Sales)))

cor(select(cigarettes, -c(State, Sales)))


summary(lm(Sales ~ Price + Income + Age, data = cigarettes))

summary(lm(Sales ~ Price + Income, data = cigarettes))

test <- cigarettes

test <- filter(cigarettes, Sales < 135 & Sales > 82)

ggplot(test, aes(x=log(Sales), y=log(Income))) + geom_point()
ggplot(test, aes(x=log(Sales), y=log(Age))) + geom_point()
ggplot(test, aes(x=log(Sales), y=log(Price))) + geom_point()

ggplot(test, aes(x=log(Sales), y=Income)) + geom_point() + 
  geom_smooth(method = "lm")
ggplot(test, aes(x=log(Sales), y=Age)) + geom_point() + 
  geom_smooth(method = "lm")
ggplot(test, aes(x=log(Sales), y=log(Price))) + geom_point() + 
  geom_smooth(method = "lm")

northeast <- c("ME", "NH", "VT", "MA", "RI", "CT", "NY", "PA", "NJ")
midwest <- c("ND", "SD", "NE", "KS", "MN", "IA", "MO", "IL", "WI", "MI", 
             "IN", "OH")
south <- c("DE", "MD", "DC", "WV", "VA", "KY", "NC", "TN", "SC", "GA",
           "AL","MS","AR","LA","OK","TX")
west <- c("WA", "OR", "ID", "MT", "WY", "CA", "NV", "UT", "CO", "AZ", "NM")

regionlookup <- rbind(data.frame(Region = "Northeast", State = northeast),
                      data.frame(Region = "Midwest", State = midwest),
                      data.frame(Region = "South", State = south),
                      data.frame(Region = "West", State = west))

cigarettesregion <- merge(cigarettes,regionlookup, by="State")

summary(lm(Sales ~ Price + Income + Age + Region, data = cigarettesregion))

summary(lm(log(Sales) ~ log(Price) + Income + Age, data = cigarettes))
summary(lm(log(Sales) ~ log(Price) + Income, data = cigarettes))
summary(lm(log(Sales) ~ log(Price) + Age, data = cigarettes))
summary(lm(log(Sales) ~ Income + Age, data = cigarettes))


```

## Data Exploration

The cigarettes data set consists of five variables: State, Age, Income, Price, and Sales. ALl of these variables are continuous except for State which is categorical. Since there is one observation per state however, this is not useful for analysis. In order to test if there are any geographical effects in this data, I divided the states into the four major census regions: Northeast, Midwest, South, and West. With multiple observations of each factor, I can generate three dummy variables to see if region has any effect.

First, I will look at a few boxplots to get a handle on the numerical variables:

```{r echo=False}

```

```{r echo=False}

# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

d1 <- ggplot(cigarettes, aes(x=1, y=Sales)) + geom_boxplot()
d2 <- ggplot(cigarettes, aes(x=1, y=Age)) + geom_boxplot()
d3 <- ggplot(cigarettes, aes(x=1, y=Income)) + geom_boxplot()
d4 <- ggplot(cigarettes, aes(x=1, y=Price)) + geom_boxplot()

multiplot(d1, d2, d3, d4, cols=2)

qplot(y=cigarettes$Sales, x= 1, geom = "boxplot")

g1 <- ggplot(cigarettes, aes(y = Sales, x = Age)) + geom_point() + 
  geom_smooth(method = "lm")
g2 <- ggplot(cigarettes, aes(y = Sales, x = Income)) + geom_point() + 
  geom_smooth(method = "lm")
g3 <- ggplot(cigarettes, aes(y = Sales, x = Price)) + geom_point() + 
  geom_smooth(method = "lm")

multiplot(g1, g2, g3, cols=2)

test <- filter(cigarettes, Sales < 135 & Sales > 82)

g1 <- ggplot(test, aes(y = Sales, x = Age)) + geom_point() + 
  geom_smooth(method = "lm")
g2 <- ggplot(test, aes(y = Sales, x = Income)) + geom_point() + 
  geom_smooth(method = "lm")
g3 <- ggplot(test, aes(y = Sales, x = Price)) + geom_point() + 
  geom_smooth(method = "lm")

g <- multiplot(g1, g2, g3, cols=2)

library(plotly)
py <- plot_ly(username='charleyferrari', key='oksysax3g0')

ggplotly(g1)

ggplot(cigarettes, aes(y = Sales)) + geom_boxplot()


install.packages("MASS")
install.packages("ISLR")
library(MASS)
library(ISLR)

Carseatdata <- Carseats[,!(names(Carseats) %in% c("Urban", "US", "ShelveLoc"))]

cormat <- cor(Carseatdata)

```