---
title: "CFerrari_Assignment11"
author: "Charley Ferrari"
date: "Monday, April 13, 2015"
output: pdf_document
---

Using R's lm function, perform regression analysis and measure the signi???cance of the independent
variables for the following two data sets. In the ???rst case, you are evaluating the statement that 
we hear that Maximum Heart Rate of a person is related to their age by the following equation:

$$ Max H R = 220 - Age $$

You have been given the following sample: 

Age   18  23  25  35  65  54  34  56  72  19  23  42  18  39  37 
MaxHR 202 186 187 180 156 169 174 172 153 199 193 174 198 183 178

Perform a linear regression analysis ???tting the Max Heart Rate to Age using the lm function in R. 
What is the resulting equation? Is the e???ect of Age on Max HR signi???cant? What is the signi???cance 
level? Please also plot the ???tted relationship between Max HR and Age. 

```{r}

age <- c(18, 23, 25, 35, 65, 54, 34, 56, 72, 19, 23, 42, 18, 39, 37)
maxHR <- c(202, 186, 187, 180, 156, 169, 174, 172, 153, 199, 193,  174, 198, 183, 178)

data <- data.frame(Age = age, MaxHR = maxHR)

summary(lm(MaxHR ~ Age, data=data))

library(ggplot2)

ggplot(data, aes(x=Age, y=MaxHR)) + 
  geom_point() +
  stat_smooth(method=lm)

```

Using the Coefficient values, the resulting eqution is:

$$MaxHR = 210.04846 - 0.79773 * Age$$

The effect of Age on MaxHR is indeed significant, satisfying a p level for the t statistic 
of 0.001. Similarly, the p value fo the F statistic is below this (on the order of magnitude of
$10^-8$.)


Using the Auto data set from Assignment 5 (also attached here) perform a Linear Regression analysis 
using mpg as the dependent variable and the other 4 (displacement, horsepower, weight, acceleration) 
as independent variables. What is the ???nal linear regression ???t equation? Which of the 4 independent 
variables have a signi???cant impact on mpg? What are their corresponding signi???cance levels? What are 
the standard errors on each of the coe???cients? Please perform this experiment in two ways. First 
take any random 40 data points from the entire auto data sample and perform the linear regression ???t 
and measure the 95% con???dence intervals. Then, take the entire data set (all 392 points) and perform 
linear regression and measure the 95% con???dence intervals. Please report the resulting ???t equation, 
their signi???cance values and con???dence intervals for each of the two runs.

```{r}

setwd("E:/Downloads/Courses/CUNY/SPS/Git/IS 605 Fundamentals of Computational Mathematics/Assignment 11/assign11")

autodata <- scan("auto-mpg.data")
autodata <- t(matrix(autodata, nrow = 5))

autodata <- data.frame(displacement = autodata[,1],
                       horsepower = autodata[,2],
                       weight = autodata[,3],
                       acceleration = autodata[,4],
                       mpg = autodata[,5])

summary(lm(mpg ~ displacement + horsepower + weight + acceleration, data=autodata))

```

Which of the 4 independent  variables have a signi???cant impact on mpg? What are their corresponding 
signi???cance levels? 

Horsepower and weight have a significant impact on mpg, satisfying p values on their t stats of
0.01 and 0.001 respectively. Displacement and acceleration don't have 

What are  the standard errors on each of the coe???cients? Please perform this 
experiment in two ways. First take any random 40 data points from the entire auto data sample and 
perform the linear regression ???t and measure the 95% con???dence intervals. Then, take the entire data 
set (all 392 points) and perform linear regression and measure the 95% con???dence intervals. Please 
report the resulting ???t equation, their signi???cance values and con???dence intervals for each of the 
two runs.

```{r}
set.seed(42)
autodataSample <- autodata[sample(1:length(autodata$displacement), 40, replace=F),]

summary(lm(mpg ~ displacement + horsepower + weight + acceleration, data=autodata))

```

The equation (when first ran) is:

$$mpg = 45.2511397 - 0.0060009*displacement - 0.0436077*horsepower - 0.0052805*weight - 0.0231480*acceleration $$

When I set the seed to 42, I actually got different significance levels:

$$mpg = 45.2511397 - 0.0060009*displacement - 0.0436077*horsepower - 0.0052805*weight - 0.0231480*acceleration $$

This will probably chan

the confidence values for the t stats are in the pr(>|t|) column in the variable list.

```{r}

summary(lm(mpg ~ displacement + horsepower + weight + acceleration, data=autodataSample))


```

For this random set, the equation is:

$$mpg = 43.264219 - 0.030943*displacement + 0.001819*horsepower - 0.004096*weight - 0.070219 *acceleration.$$

This time