---
title: "608SemesterProject"
author: "Charley Ferrari"
date: "Tuesday, April 14, 2015"
output: pdf_document
---

First, lets read the data in and libraries.

```{r}

library(ggplot2)
library(ggthemes)
library(dplyr)

setwd("E:/Downloads/Courses/CUNY/SPS/Git/IS 608 Knowledge and Visual Analytics/Semester Project")

agent <- read.csv("agentdataprop.csv")

econ <- read.csv("econdataprop.csv")

econ$Date <- strptime(econ$Date, format="%m/%d/%Y")



```

I've chosen to use the company visit scores as my base data. On page 64 of the [supporting documentation](http://www.bankofengland.co.uk/publications/Documents/quarterlybulletin/2013/qb130106.pdf), seven of the eleven scores are mapped to macroeconomic concepts: 

**Demand and Output**

Total Demand : Private Sector Nominal Output

Exports : Exports

Investment : Business Investment

**Factor Utilisation**

Employment :  Private Sector Unemploymen

**Costs and Prices**

Total Labour Costs : Average Weekly Earnings Total Pay

Pay : Average Weekly Earnings Regular Pay

Profits : Gross Operating Surplus

My visualization will start out with two line graphs: On the top, a graph of the economic concept,
and below, the aggregate company visit scores



```{r}

agent01 <- agent %>%
  select(ActualDateDisplay, DemandScore) %>%
  group_by(ActualDateDisplay) %>%
  summarize(MeanCVS = mean(DemandScore))

agent01$ActualDateDisplay <- strptime(agent01$ActualDateDisplay, format="%m/%d/%Y")

agent01 <- agent01[order(agent01$ActualDateDisplay),]

ggplot(econ, aes(x=Date, y=ValueAdd)) +
  geom_line() + 
  theme_tufte()


ggplot(agent01, aes(x=ActualDateDisplay, y=MeanCVS)) +
  geom_line() +
  theme_tufte()


```

At this stage, users would be able to choose their concept from a drop down menu. If they wanted
to select exports for example:

```{r}

agent02 <- agent %>%
  select(ActualDateDisplay, ExportScore) %>%
  group_by(ActualDateDisplay) %>%
  summarize(MeanCVS = mean(ExportScore))

agent02$ActualDateDisplay <- strptime(agent02$ActualDateDisplay, format="%m/%d/%Y")

agent02 <- agent02[order(agent02$ActualDateDisplay),]

ggplot(econ, aes(x=Date, y=Exports)) +
  geom_line() + 
  theme_tufte()


ggplot(agent02, aes(x=ActualDateDisplay, y=MeanCVS)) +
  geom_line() +
  theme_tufte()



```

Instead of displaying the distribution of the CVS as a pop up, I think it would make more sense
to display it as a third panel below these two graphs. I'd ideally like to have the graph show
values when the user hovers over them, and allow the user to select a specifc time period by
clicking on the point being hovered over.

So, lets say the user clicks on 01/01/2010:

```{r}

agent03 <- agent %>%
  filter(ActualDateDisplay == "1/1/2010") %>%
  select(Sector, ExportScore)

ggplot(econ, aes(x=Date, y=Exports)) +
  geom_line() + 
  theme_tufte()


ggplot(agent02, aes(x=ActualDateDisplay, y=MeanCVS)) +
  geom_line() +
  theme_tufte()

ggplot(agent03, aes(x=ExportScore)) +
  geom_density(fill="blue", alpha=.2) +
  theme_tufte()
  


```

Now, in this third pane, I'd like to allow the user to select a sector via a drop down menu.
Once the user selects a sector, the distribution will change, and the average CVS in pane 2
will change.

I'm still not sure how I will tweak the economic concepts. Some roll up nicely to the industries
(such as employment), while others don't. Right now I'll just assume the economic concept remains
the same, but I could just as easily imagine the selection of a sector will also change the 
headline economic concept

So, lets say the user chose "Business and Financial Services"

```{r}

agent04 <- agent %>%
  filter(Sector == "Business and financial services") %>%
  select(ActualDateDisplay, ExportScore) %>%
  group_by(ActualDateDisplay) %>%
  summarise(MeanCVS = mean(ExportScore))

agent04$ActualDateDisplay <- strptime(agent04$ActualDateDisplay, format="%m/%d/%Y")

agent04 <- agent04[order(agent04$ActualDateDisplay),]

agent05 <- agent03 %>% 
  filter(Sector == "Business and financial services")

ggplot(econ, aes(x=Date, y=Exports)) +
  geom_line() + 
  theme_tufte()


ggplot(agent04, aes(x=ActualDateDisplay, y=MeanCVS)) +
  geom_line() +
  theme_tufte()

ggplot(agent05, aes(x=ExportScore)) +
  geom_density(fill="blue", alpha=.2) +
  theme_tufte()

```

There are a few other selections I could allow the user to make. Future scores are given in 
addition to backward looking scores, so that could just as easily be added in a drop down menu
that affects all panes.

In reading through the documentation, I had another insight into how this data can be viewed:
the distributions can be animated. Chart 4 shows the distribution of CVS profit scores in 2009
versus 2012. I think this could be viewed as an animation, showing how the density curve changes
through time.

More interestingly, charts 5 and 8 show scatter plots just begging to be changed into Hans 
Rosling style animated charts. 

In pane 3, I'd like to add a "play" button, that would allow the distribution to be animated, and
return to the date chosen in pane 1 or 2. I'd also like to add a 4th pane that would allow the 
user to pick another concept to compare. This would allow the sector picked in pane 3 to apply
to the scatter plot.





